### 2019-05-02 papers/haskell19/6-casestudies.tex
Added automatic representaition balancing

We benchmarked the execution time required to generate and fully evaluate 1000
random values corresponding to each case study, comparing both manually-written
concrete generators and modular generators obtained using our approach.
%
For this purpose, we used the \emph{Criterion} benchmarking tool for Haskell,
and limited the maximum depth of the generated values to five levels.
%
Additionally, our modular generators were tested using both linear and balanced
generation specifications.
%
Figure \ref{fig:times} illustrates the relative execution time of each case
study, with respect to each manually-written random generator---we encourage the
reader to obtain a colored version of this work.

\begin{figure}[b]
  \centering
  \begin{tikzpicture}
    \begin{axis}
      [ height=6cm
      , width=9cm
      , ybar
      , ymajorgrids=true
      , ytick={0,1,...,5}
      , ymax=4.5
      , enlargelimits=0.25
      , enlarge x limits=0.25
      , /pgf/bar width=17pt
      , legend style={
        at={(0.5,-0.15)},
        anchor=north,legend columns=3}
      , legend image code/.code={%
        \draw[#1, draw=none] (0cm,-0.1cm) rectangle (0.3cm,0.2cm); }
      , ylabel={Relative Generation Time}
      , symbolic x coords={RBT, SExp, HTML}
      , xtick=data
      % , nodes near coords
      , nodes near coords align={vertical}
      ]
      \addplot[fill, red!30]   coordinates {(SExp,1)    (RBT,1)    (HTML,1)};
      \addplot[fill, green!30] coordinates {(SExp,2.31) (RBT,3)    (HTML,4.88)};
      \addplot[fill, blue!30]  coordinates {(SExp,1.62) (RBT,2.28) (HTML,1.70)};
      \legend{Manual, Modular/Linear, Modular/Balanced}
    \end{axis}
  \end{tikzpicture}
  \caption{Generation time comparison between concrete and modular generators.}
  \label{fig:times}
\end{figure}

As it can be observed, our approach suffers from a noticeable runtime overhead,
specially when considering the HTML case study with a large amount of
construction involved at the generation process.
%
However, by balancing our representation types we found that the generation
performance gets dramatically improved.
%
To apply this optimization automatically, and relying on more sophisticated
techniques like prefix-free encodings to obtain optimal representations, is a
undergoing path of our future work.

@@

Nonetheless, we consider that these results are in turn very encouraging, given
that the flexibility obtained from using our approach does not produce severe
slowdowns in practice, nor explodes more than linearly as we increase the
complexity of our random generators.

>>>>

We benchmarked the execution time required to generate and fully evaluate 10000
random values corresponding to each case study, comparing both manually-written
concrete generators and modular generators obtained using our approach.
%
For this purpose, we used the \emph{Criterion} benchmarking tool for Haskell,
and limited the maximum depth of the generated values to five levels.
%
Additionally, our modular generators were tested using both linear and balanced
generation specifications.
%
Figure \ref{fig:times} illustrates the relative execution time of each case
study, with respect to each manually-written random generator---we encourage the
reader to obtain a colored version of this work.

\begin{figure}[b]
  \centering
  \begin{tikzpicture}
    \begin{axis}
      [ height=6cm
      , width=9cm
      , ybar
      , ymajorgrids=true
      , ytick={0,1,...,4}
      , ymax=4
      , ymin=0.5
      , enlargelimits=0.25
      , enlarge x limits=0.25
      , /pgf/bar width=17pt
      , legend style={
        at={(0.5,-0.15)},
        anchor=north,legend columns=3}
      , legend image code/.code={%
        \draw[#1, draw=none] (0cm,-0.1cm) rectangle (0.3cm,0.2cm); }
      , ylabel={Relative Generation Time}
      , symbolic x coords={RBT, SExp, HTML}
      , xtick=data
      % , nodes near coords
      , nodes near coords align={vertical}
      ]
      \addplot[fill, red!30]   coordinates {(SExp,1)    (RBT,1)    (HTML,1)};
      \addplot[fill, green!30] coordinates {(SExp,2.40) (RBT,3.57) (HTML,4.27)};
      \addplot[fill, blue!30]  coordinates {(SExp,1.60) (RBT,2.52) (HTML,1.55)};
      \legend{Manual, Modular (lin), Modular (bal)}
    \end{axis}
  \end{tikzpicture}
  \caption{Generation time comparison between concrete and modular generators.}
  \label{fig:times}
\end{figure}

As it can be observed, our approach suffers from a noticeable runtime overhead
when using linearly defined representations, specially when considering the HTML
case study with a large amount of construction involved at the generation
process.
%
However, we found that, by balancing our representation types the generation
performance improves dramatically.
%
At the light of these improvements in the performance, we included in our tool
an additional type-level computation that automatically balances our
representations in order to reduce the generation overhead.

% and relying on more sophisticated techniques like prefix-free encodings to
% obtain optimal representations, is a undergoing path of our future work.

@@

All in all, we consider that these results are fairly encouraging, given that
the flexibility obtained from using our compositional approach does not produce
severe slowdowns in whem generating random values in practice.

% , nor explodes more than linearly as we increase the complexity of our random
% generators.
